<span style="color:red;">*ANTs*</span><br> <small>_power tools for neuroimaging_</small>

### 2014-10-23 ants2015_GE.Rmd
ENH:  More edits to the talk.

@@

# our strong <span style="color:red;">mathematical and software engineering</span> foundation leads to near limitless opportunities for innovation in a variety of application domains

>>>>

<!--

@@

# our strong <span style="color:red;">mathematical and software engineering</span> foundation leads to near limitless opportunities for innovation in a variety of application domains
-->

### 2014-10-23 ants2015build.Rmd
ENH:  More edits to the talk.

# our strong <span style="color:red;">mathematical and software engineering</span> foundation leads to near limitless opportunities for innovation in a variety of application domains

>>>>

# our strong <span style="color:red;">mathematical and software engineering</span> foundation leads to near limitless opportunities for innovation in a variety of application domains
-->

# ANTs brain morphology<br><br>[<small>"*Large-scale evaluation of ANTs and FreeSurfer cortical thickness measurements*"</small>](http://www.ncbi.nlm.nih.gov/pubmed/24879923)

## Cortical thickness studies:
```{r, results='asis', echo = FALSE}
conditions <- c( "Huntington’s disease", "schizophrenia", "bipolar disorder",
                 "Alzheimer’s disease", "frontotemporal dementia", "Parkinson’s disease",
                 "Williams syndrome", "multiple sclerosis", "autism", "migraines",
                 "chronic smoking", "alcoholism", "cocaine addiction", "Tourette syndrome in children",
                 "scoliosis in female adolescents", "early-onset blindness", "chronic pancreatitis",
                 "obsessive-compulsive disorder", "ADHD", "obesity", "heritable depression",
                 "elderly depression", "age", "gender", "untreated male-to-female transsexuality",
                 "handedness", "intelligence", "athletic ability", "meditative practices", "musical ability",
                 "tendency toward criminality", "childhood sexual abuse in female adolescents",
                 "Tetris-playing ability in female adolescents", "traumatic brain injury" )

conditionsDF <- data.frame( Column1 = conditions[1:floor( 0.5*length(conditions) )], Column2 = conditions[( 1+floor( 0.5*length(conditions) ) ):length(conditions)] )
knitr::kable( conditionsDF )
```

## *ANTs*-related background

* Sandy, Brian, and <span style="color:red;">DiReCT</span>
  + used other tools such as FMRIB's Automated Segmentation Tool (FAST)
  + `KellySlater.cxx`
  * scripts "adequate for Penn data"
* `KellyKapowski.cxx` and `antsCorticalThickness.sh`
* _"Let's evaluate on open data."_
* and, eventually, _"Let's compare with FreeSurfer."_

## The *ANTs* structural brain mapping workflow

![](./evaluation/figures/pipeline.png)

## Basic components of the pipeline

<div align="center"><img height="350" src="./evaluation/figures/components.png" frameborder="0"></img></div>

1. template building (offline)
2. brain extraction
3. cortical thickness estimation
4. cortical parcellation

## Template building

*Tailor data to your specific cohort*

<div align="center"><img height="350" src="./evaluation/figures/templates.png" frameborder="0"></img></div>

* Templates representing the average mean shape and intensity are built directly from the cohort to be analyzed, e.g. pediatric vs. middle-aged brains.
* Acquisition and anonymization (e.g. defacing) protocols are often different.

## Template building (cont.)

<div align="center"><img height="350" src="./evaluation/figures/templatePriors.png" frameborder="0"></img></div>

Each template is [processed](https://github.com/ntustison/antsCookTemplatePriorsExample)
to produce auxiliary images which are used for brain extraction and brain segmentation.

## Brain extraction comparison<br>*ANTs* vs. FreeSurfer

<div align="center"><img height="380" src="./evaluation/figures/brainExtraction.png" frameborder="0"></img></div>

Comparison with de facto standard FreeSurfer package.  Note the difference in separation of
the gray matter from the surrounding CSF.  (0 failures out of 1205 scans)

## Brain segmentation

<div align="center"><img height="350" src="./evaluation/figures/brainSegmentation.png" frameborder="0"></img></div>

Randomly selected healthy individuals.  Atropos gets good performance across ages.

## Cortical thickness estimation

<div align="center"><img height="250" src="./evaluation/figures/corticalThicknessEstimation.png" frameborder="0"></img></div>

In contrast to FreeSurfer which warps coupled surface meshes to segment the gray matter, *ANTs* diffeomorphically registers the white matter to the combined gray/white matters while simultaneously estimating thickness.

## _But without ground truth, how does one evaluate the pipeline?_

* Use age and gender (demographic information) in training/prediction paradigm
* Public data sets (IXI, NKI, Oasis, MMRR)
* $>$ 1200 subjects, age 7 to over 90 years old

## Prediction from cortical thickness data:<br>gender and age

<div align="center"><img height="350" src="./evaluation/figures/evaluation.png" frameborder="0"></img></div>

## Regional importance comparison

$AGE \sim VOLUME + GENDER + \sum_{i=1}^{62} T(DKT_i)$

<div align="center"><img height="500" src="./evaluation/figures/antsvfreesurfer_importance.png" frameborder="0"></img></div>

*ANTs* (left) vs. FreeSurfer (right)

## Regional importance comparison (cont.)

<div align="center"><img height="550" src="./evaluation/figures/antsvfreesurfer_regionalPlots.png" frameborder="0"></img></div>

## Cortical thickness extensions
